---
title: "The use of eDNA to assess remote coastal islands for the presence of invasive terrestrial mammals"
author: "Doug Hynes"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(taxize)
library(data.tree)
```

### **Abstract**

Coastal islands face grave threats such as sea level rise, development, and biological invasions. Yet, in many areas, basic knowledge of the prevalence of island dwelling species and the pressures that limit their persistence is often lacking. Biological inventories can help fill these knowledge gaps and inform converation actions, but are hindered by the logistics and costs of conducting offshore or nearshore fieldwork. In light of the need for environmental practitioners to be able to survey islands robustly and efficiently, we test off-the-shelf commmerical environmental DNA water samplers and commerical metabarcoding lab services as ways to enable rapid inventories of coastal islands for mammalian biodiversity and invasive terrestrial threats.

To date, reads (i.e., sequences of base pairs that correspond to part of a DNA fragment) that correspond with 45 taxon have been detected from the 35 water samples (collected from islands Mud, Flat, Seal, and Boot). 


```{r message = FALSE, echo=FALSE, comment=NA}

#cleanData <- read_csv("C:/Users/HynesD/eDnaTools/cleanedData.csv")

# getNames <-
#   tax_name(
#     sumData2$taxa,
#     get = c(
#       "phylum",
#       "class",
#       "order",
#       "suborder",
#       "superfamily",
#       "family",
#       "subfamily",
#       "genus",
#       "species"
#     ),
#     db = "ncbi"
#   ) %>%
#   rename(taxa = "query") %>% replace_na(list(class = "Reptilia")) %>%
#   inner_join(sumData2)
# 
# getNames$pathString <-
#   paste(
#     getNames$phylum,
#     getNames$class,
#     getNames$order,
#     getNames$suborder,
#     getNames$superfamily,
#     getNames$family,
#     getNames$subfamily,
#     getNames$genus,
#     getNames$species,
#     sep = "/"
#   )
# 
# getNames$pathString <- ""
# groups <-
#   c(
#     "phylum",
#     "class",
#     "order",
#     "suborder",
#     "superfamily",
#     "family",
#     "subfamily",
#     "genus",
#     "species"
#   )
# 
# for (i in 1:nrow(getNames)) {
#   isNA <- TRUE
#   
#   for (j in 11:3) {
#     if (isNA) {
#       if (!is.na(getNames[i, j])) {
#         getNames$pathString[i] <-
#           paste(getNames[i, j], getNames$pathString[i], sep = "/")
#         isNA <- FALSE
#       }
#     } else {
#       getNames$pathString[i] <-
#         paste(getNames[i, j], getNames$pathString[i], sep = "/")
#     }
#   }
# }

#write_csv(getNames, "C://Users/HynesD/eDnaTools/data/taxonNames.csv")
getNames <- read_csv("C://Users/HynesD/eDnaTools/data/taxonNames.csv")

rodents <- getNames %>% filter(order %in% c("Rodentia"))
tree <- as.Node(rodents)
#rodents <- tree$Rodentia
#Sort(northAm, "GNI", decreasing = TRUE)
#print(rodents, "readsT")
print(tree, "readsT", pruneMethod = "simple")
```




Muroidea - a superfamily consisting of rats, mice, voles  
Murinae - subfamily old world rats and mice  
Cricetidae - new world rats and mice, voles (e.g. *Myodes*, *Microtus*, *Lemmus*, *Peromyscus*)  
Arvicolinae - subfamily, includes muskrat  

Lagamorpha - i.e., Lepus americanus  
Leporidae i.e., Lepus americanus  
Lepus

Sus - domestic pig  

Soricidae - Probably represents masked shrew; less probable smokey shrew or others  
Soricinae - sunfamily, includes *Sorex* (e.g., *Blarina*)  
Sorex - genus includes smoky shrew, water shrew, etc.  

Leuciscidae - true minnows  
Fundulus - more than one species of *Fundulus* possible across study extent

